---
title: "Hilary_variable_cleaning"
author: "Hilary McLeland-Wieser"
date: "2023-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#setup and loading in data
rm(list=ls())
if (!require(pacman)) install.packages('pacman', repos = 'https://cloud.r-project.org')
#REDCap generated code to add labels to all variables, make relevant variables factors, and assign levels
pacman::p_load(dplyr, tidyr, janitor, ggplot2, readr, tidyverse, lubridate, naniar, Hmisc, gmodels)
load("/projects/prlab/unrestricted/Center Projects/Healthy Dairy Worker/Data and Analysis/Data/cleaned_1.1.23.Rdata")
```

New Variable: growup_on_farm_01, growup_on_farm_fact
Variables Used: growup_on_farm___1, growup_on_farm___2, growup_on_farm_v3
Cleaning: For growup_on_farm___1 and growup_on_farm___2 change 1 to Yes/No and 0 to NA. Combine all three gorwuo_on_fam variables into new logical and factor variables. 
Class: growup_on_farm_01 = logical, growup_on_farm_fact = factor
Description: Joint variables for "Did you grow up on a farm ages birth-10?"

```{r, Grew up on Farm}
table(data$growup_on_farm___1, data$growup_on_farm___2, deparse.level=2, useNA = "always")
table(data$growup_on_farm_v3, deparse.level=2, useNA = "always")

data$growup_on_farm___1[data$growup_on_farm___1==1] <- "Yes"
data$growup_on_farm___1[data$growup_on_farm___1==0] <- NA

data$growup_on_farm___2[data$growup_on_farm___2==1] <- "No"
data$growup_on_farm___2[data$growup_on_farm___2==0] <- NA

data <- data %>% mutate(growup_on_farm_01 = case_when(growup_on_farm___1 == "Yes" ~ 1,
                                           growup_on_farm___2 == "No" ~ 0,
                                           growup_on_farm_v3 == 1 ~ 1,
                                           growup_on_farm_v3 == 2 ~ 0))

data$growup_on_farm_fact <- factor(data$growup_on_farm_01, levels = 0:1, labels = c("No", "Yes"))

table(data$growup_on_farm_fact, data$growup_on_farm_01, useNA = "always")
```

